\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-07-23 doc/sw-system.tex
Updated purity numbers and discussion as suggested by Anu, folded in Aprajita's comments on selection function text. Finished response with promise of latexdiff'ed PDF... #216

The purity depends on the proportion of sims to
duds, and so the purity of the test set must be approximated by rescaling the
training set to the expected proportion of lens systems to not-lens systems in
the survey. First we compute the expected number of false positives by
multiplying the FPR by the expected number of non-lenses in the survey. Then
we multiply the TPR by the expected number of lenses in the survey, to get the
expected number of true positives. The sum of the true positives and the false
positives gives the expected sample size; dividing the expected number of true
positives by this sample size gives the purity. Note that the completeness is
invariant to this transformation.
The \StageOne curves are truncated by the retirement of subjects in this phase,
which sets the minimum size of this sample. We see from the solid blue curve
that over 90\% completeness was reached, albeit in a sample with 30\%
purity.

@@

Finally, it is worth noting the implications of the results in this
section for future studies of samples of lenses (and lens candidates)
discovered through visual inspection at \SW. The ROC curve analysis we
have carried out should be very familiar to those working in machine
classification, and in fact is identical to that which would be
performed on the classifications made by a new automated method.
Supervised machine learning methods and the \SW system as described
here both require a training set, and both return quantitative,
probabilistic classifications (consisting of a label and some measure
of confidence in that label) that can be used in further analysis. A
good example of where such quantitification is important is in the
derivation of the selection function for a given sample. While such a
derivation is beyond the scope of this work, it is sufficient to note
that from the point of view of an astronomer seeking to derive a
selection function, {\it the labels produced by \SW and those produced
by a machine learning method can be treated equivalently}: that is, we
have succeeded in elevating visual inspection to a quantitative
science. Indeed, the discussion of completeness given above is all in
terms of recovery of a large set of simulated lenses, just as it would
be in the case of an automated method---and in both cases, the
limiting factor is the realism of the training set. In \PaperTwo we
investigate this further by assessing the performance of the \SW
system against the (small) set of real lens candidates known to lie in
the field; for now, we note that the training set used in this work
constitutes a valid sample of objects for any alternative machine
learning lens detection method to be tested against. In future, we may
well find it necessary to employ a larger training set, to enable the
selection efficiency to be characterised as a function of multiple
observables. This should not be a problem, since in the current work
the small training set we used was classified around 20 times more
than  was the test set: we could have defined a larger training set,
and collected  many fewer classifications per training image.

@@

\item The optimal true positive rate (completeness) and false positive rate in
the training set were estimated to be around $91-93\%$ and $<1\%$
(in both classification stages).
This FPR translates to a purity of approximately 30\%. We find an even
higher purity at \StageTwo (perhaps as a result of the least visible candidates already
being discarded): here, 100\% purity was reached at just under $90\%$
completeness. Because \SW is a supervised learning system, these numbers should
be taken to be upper limits to what we should expect for real lenses (which have
not been selected for a high visibility training set).

>>>>

The purity depends on the proportion of sims to duds, and so the
purity of the test set must be approximated by rescaling the training
set to the expected proportion of lens systems to not-lens systems in
the survey. We expect there to be around 90 lenses in the CFHTLS
already (a rate of 1 lens in every 5000 images or so); to a very good
approximation  the number of non-lens images in the survey is just the
number of images in  the survey (some 430,000). First we compute the
expected number of false positives by multiplying the FPR by the
expected number of non-lenses in the survey (430,000).  Then we
multiply the TPR by the expected number of lenses in the survey (90),
to get the expected number of true positives. The sum of the true
positives and the false positives gives the expected sample size;
dividing the expected number of true positives by this sample size
gives the purity. Note that the completeness is invariant to this
transformation. The \StageOne curves are truncated by the retirement
of subjects in this phase, which sets the minimum size of this sample.
We see from the solid blue curve that over 90\% completeness was
able to be reached, albeit in samples with not more than 30\% purity.
We set the detection threshold for \StageOne to be 0.95 (shown by the blue star
in \Fref{fig:results:sample:CP}), leading to a sample with 94\% completeness
and 15\% purity.

@@

Finally, it is worth noting the implications of the results in this
section for future studies of samples of lenses (and lens candidates)
discovered through visual inspection at \SW. The ROC curve analysis we
have carried out should be very familiar to those working in machine
classification, and in fact is identical to that which would be
performed on the classifications made by a new automated method.
Supervised machine learning methods and the \SW system as described
here both require a training set, and both return quantitative,
probabilistic classifications (consisting of a ``label'' and some measure
of confidence in that label) that can be used in further analysis. A
good example of where such quantitification is important is in the
derivation of the selection function for a given sample. While such a
derivation is beyond the scope of this work, it is sufficient to note
that from the point of view of an astronomer seeking to derive a
selection function, {\it the labels produced by \SW and those produced
by a machine learning method can be treated equivalently}: that is, we
have succeeded in elevating visual inspection to a quantitative
science. Indeed, the discussion of completeness given above is all in
terms of recovery of a large set of simulated lenses, just as it would
be in the case of an automated method---and in both cases, the
limiting factor is the realism of the training set. In \PaperTwo we
investigate this limit further by assessing the performance of the \SW
system against the (small) set of real lens candidates known to lie in
the field; for now, we note that the training set used in this work
constitutes a valid sample of objects for any alternative machine
learning lens detection method to be tested against. In future, we
could employ a larger training set, to enable the
selection efficiency to be characterised as a function of multiple
observables. This will be achievable, since
the small training set we used in the current study
was classified around 20 times more
than  was the test set: we could therefore collect many fewer
classifications per training image in a much larger training set.

@@

\item The optimal true positive rate (completeness) and false positive rate in
the training set were estimated to be around $92-94\%$ and $<1\%$
(in both classification stages).
This FPR translates to a purity of approximately $15-30$\%. We find that even
higher purities were achievable at \StageTwo
(perhaps as a result of the least visible candidates already
being discarded): here, 100\% purity was reached at just under $90\%$
completeness. Because \SW is a supervised learning system, these numbers should
be taken to be upper limits to what we should expect for real lenses (which have
not been selected for a high visibility training set).
